Locality and Computational Reliability of Linear Response Calculations for Molecular Systems

### 2018-07-12 AnalyticStructure_main.tex
update

Such locality property of the fluctuation state can be directly confirmed for static perturbations at $\omega=0$.
Let us consider a static perturbing operator $\op\Phi_s$ that can be restricted to a finite domain. 
Any local operator, like \emph{e.g.} a static electric field $\mathbf F$, with associated scalar potential $\Phi_s[\mathbf F](\r) = -\mathbf F \cdot \r$, would satisfy such property.
We consider the perturbed Hamiltonian $\hnot[\dm_{\Phi_s}] + \op\Phi_s[\mathbf F]$, where $\dm_{\Phi_s}$ represents the GS density of the perturbed system, and assume that the 
perturbation strength ($|\mathbf F|$ in the above example) is sufficiently small to be consistent with a LR treatment of the problem. 
In this case, the occupied eigenfunction $\ket{\psi_p^{\Phi_s}}$ of the perturbed Hamiltonian can be extracted with traditional GS techniques, 
as from the discussion around Eq.~\eqref{SEHelmholtz1}, they exhibit bound-state behavior.
\begin{figure}[t]
\includegraphics[scale=0.68]{CO_statPolvsBox.eps}
\caption{\label{co_alphaStatic}(Color online) Convergence of the statical polarizability $\alpha_{zz}$ (blue continuous line) and of the ground state dipole $d_z$ (red dotted line) 
of $CO$ molecule as a function of the size of the computational domain. The inset evidences the relative convergence (in percent) of the two quantities w.r.t. the corresponding value associated to the largest 
box.}
\end{figure}

@@

On the contrary, the computation of the linear response functional can be much more demanding in the \emph{above threshold regime}, realized for $\omega>|\eps_h|$, in which fluctuation 
states start behaving as unbound wave functions.
As by hypothesis $\ket{\op O \psi_p}$ is bound-state like, \emph{only} the scalar products of Eq.~\eqref{LinearResponseFunctDef2} can be evaluated in a localized domain.
In this case a computational description of fluctuation states is only meaningful in ``weak'' sense,
%\emph{i.e.} inside the scalar products of 
%\eqref{LinearResponseFunctDef2}. 
%since $\ket{f_p^\Phi(\omega)}$ becomes an unbound, delocalized function, 
where the bound state character of $\{\bra{\psi_p}\op O\}$ behaves as a regulator.

@@

\begin{figure}
\includegraphics[scale=0.6]{CO_spectrum.eps}
\caption{\label{co_spectrum}(Color online) Illustrative LR absorption spectra of the $CO$ molecule in computational setups with different sizes of the simulation domain. The value of the IP threshold is indicated by a vertical dashed line.}
\end{figure}

@@

\begin{figure}[ht]
\includegraphics[scale=0.6]{CO_dos.eps}
\caption{(Color online) DOS of $CO$ excitation energies $\Omega_a$. The two curves correspond to two computational domain of different sizes. The inset contains the sole contribution of localized excitations. The first IP threshold is depicted as the vertical dashed line.}
\label{CO_exc}
\end{figure}
\begin{figure}[ht]
\centering
\subfloat[Excitation landscape]
{\label{c6h6_excLand}
\includegraphics[scale=0.56]{C6H6_excLandscape.eps}} \\
\centering
\subfloat[DOS of Excitation energies]
{\label{C6H6_exc}
\includegraphics[scale=0.58]{C6H6_dos.eps}}
\caption{(Color online) Benzene molecule. (a): Excitation Landscape: the excitations have been separated in discrete and continuum sector following the criteria of Sec.~\ref{ExcitationLandscape}. (b): DOS of excitation energies $\Omega_a$. The first IP threshold is depicted as the vertical dashed line.}
\end{figure}

@@

\begin{figure}
\includegraphics[scale=0.6]{CO_statPolvsExc.eps}
\caption{\label{co_AlphaExc}(Color online) Convergence of the statical polarizability $\alpha_{zz}$ of the $CO$ molecule, in a box of $11.5$ \AA,  
w.r.t the excitations considered in the integral of Eq.~\eqref{sminus2}. The reference value obtained from Eq.~\eqref{staticalpha} is represented as a 
horizontal red dashed line. The value of the IP energy is indicated by a vertical black dashed line. For each excitation energy, the values of the 
oscillator strengths $w^a_{zz}$ of Eq.~\eqref{sminus2} are also plotted. }
\end{figure}
Hence, contrary to GS, \emph{for LR calculation,
even when the response density is localized, the contribution coming from the 
continuum sector of the excitations can never be neglected}. 
In other terms, we \emph{cannot} exclude that a fluctuation state, even if genuinely localized, will have a nonzero projection
in the essential part of the Liouvillian spectrum.
This fact shows that it is \emph{impossible} to restrict the spectral representation
of $\chi(\omega)$ only localized, discrete, \emph{code-independent} excitations, even for representing LR results in the $\omega=0$ limit.

>>>>

Such locality property of the fluctuation state can be directly confirmed for static perturbations at $\omega=0$.
Let us consider a static perturbing operator $\op\Phi_s$ that can be restricted to a finite domain. 
Any local operator, like \emph{e.g.} a static electric field $\mathbf F$, with associated scalar potential $\Phi_s[\mathbf F](\r) = -\mathbf F \cdot \r$, would satisfy such property.
We consider the perturbed Hamiltonian $\hnot[\dm_{\Phi_s}] + \op\Phi_s[\mathbf F]$, where $\dm_{\Phi_s}$ represents the GS density of the perturbed system, and assume that the 
perturbation strength ($|\mathbf F|$ in the above example) is sufficiently small to be consistent with a LR treatment of the problem. 
In this case, the occupied eigenfunction $\ket{\psi_p^{\Phi_s}}$ of the perturbed Hamiltonian can be extracted with traditional GS techniques, 
as from the discussion around Eq.~\eqref{SEHelmholtz1}, they exhibit bound-state behavior.
\begin{figure}[t]
\includegraphics[scale=0.68]{Fig1_CO_statPolvsBox.eps}
\caption{\label{co_alphaStatic}(Color online) Convergence of the statical polarizability $\alpha_{zz}$ (blue continuous line) and of the ground state dipole $d_z$ (red dotted line) 
of $CO$ molecule as a function of the size of the computational domain. The inset evidences the relative convergence (in percent) of the two quantities w.r.t. the corresponding value associated to the largest 
box.}
\end{figure}

@@

On the contrary, the computation of the linear response functional can be much more demanding in the \emph{above threshold regime}, realized for $\omega>|\eps_h|$, in which fluctuation 
states start behaving as unbound wave functions.
As by hypothesis $\ket{\op O \psi_p}$ is bound-state like, \emph{only} the scalar products of Eq.~\eqref{LinearResponseFunctDef2} can be evaluated in a localized domain.
In this case a computational description of fluctuation states is only meaningful in ``weak'' sense,
where the bound state character of $\{\bra{\psi_p}\op O\}$ behaves as a regulator.

@@

\begin{figure}
\includegraphics[scale=0.6]{Fig2_CO_spectrum.eps}
\caption{\label{co_spectrum}(Color online) Illustrative LR absorption spectra of the $CO$ molecule in computational setups with different sizes of the simulation domain. The value of the IP threshold is indicated by a vertical dashed line.}
\end{figure}

@@

\begin{figure}[ht]
\includegraphics[scale=0.6]{Fig4_CO_dos.eps}
\caption{(Color online) DOS of $CO$ excitation energies $\Omega_a$. The two curves correspond to two computational domain of different sizes. The inset contains the sole contribution of localized excitations. The first IP threshold is depicted as the vertical dashed line.}
\label{CO_exc}
\end{figure}
\begin{figure}[ht]
\centering
\subfloat[Excitation landscape]
{\label{c6h6_excLand}
\includegraphics[scale=0.56]{Fig5a_C6H6_excLandscape.eps}} \\
\centering
\subfloat[DOS of Excitation energies]
{\label{C6H6_exc}
\includegraphics[scale=0.58]{Fig5b_C6H6_dos.eps}}
\caption{(Color online) Benzene molecule. (a): Excitation Landscape: the excitations have been separated in discrete and continuum sector following the criteria of Sec.~\ref{ExcitationLandscape}. (b): DOS of excitation energies $\Omega_a$. The first IP threshold is depicted as the vertical dashed line.}
\end{figure}

@@

\begin{figure}
\includegraphics[scale=0.6]{Fig6_CO_statPolvsExc.eps}
\caption{\label{co_AlphaExc}(Color online) Convergence of the statical polarizability $\alpha_{zz}$ of the $CO$ molecule, in a box of $11.5$ \AA,  
w.r.t the excitations considered in the integral of Eq.~\eqref{sminus2}. The reference value obtained from Eq.~\eqref{staticalpha} is represented as a 
horizontal red dashed line. The value of the IP energy is indicated by a vertical black dashed line. For each excitation energy, the values of the 
oscillator strengths $w^a_{zz}$ of Eq.~\eqref{sminus2} are also plotted. }
\end{figure}
Hence, contrary to GS, \emph{for LR calculation,
even when the response density is localized, the contribution coming from the 
continuum sector of the excitations can never be neglected}. 
In other terms, we \emph{cannot} exclude that a fluctuation state, even if genuinely localized, will have a nonzero projection
in the essential part of the Liouvillian spectrum.
This fact shows that it is \emph{impossible} to restrict the spectral representation
of $\chi(\omega)$ only localized, discrete, \emph{code-independent} excitations, even for representing LR results in the $\omega=0$ limit.
